---
title: "PHDS Data Pipeline Report"
format:
  html:
    toc: true
    toc-depth: 2
execute:
  echo: true
  warning: false
  error: true
---

## Overview

This report summarizes the **pediatric visit dataset** produced by the phds-data-pipeline-as-code pipeline. The dataset is **synthetic**—generated with deterministic seeds for development, testing, and demos.

**Privacy note:** No real patient or clinical data is used. All records are artificially generated and do not represent any individual.

```{r setup, include=FALSE}
# Load pipeline outputs (written by targets before render)
# Robust for CI: no interactive prompts, graceful fallbacks
metrics <- if (file.exists("output/metrics.rds")) readRDS("output/metrics.rds") else list()
validation <- if (file.exists("output/validation.rds")) readRDS("output/validation.rds") else list(valid = NA, errors = character(0), warnings = character(0))
df <- if (file.exists("data/synthetic/visits.parquet")) arrow::read_parquet("data/synthetic/visits.parquet") else NULL
contract_raw <- if (file.exists("data_contract/contract.yaml") && requireNamespace("yaml", quietly = TRUE)) yaml::read_yaml("data_contract/contract.yaml") else list()
```

## Data Contract

Key rules from `data_contract/contract.yaml`:

```{r contract, results='asis'}
if (length(contract_raw) > 0) {
  meta <- contract_raw$metadata
  ds <- contract_raw$datasets$pediatric_visits
  v <- if (is.null(meta$version)) "—" else meta$version
  o <- if (is.null(meta$owner)) "—" else meta$owner
  u <- if (is.null(meta$last_updated)) "—" else meta$last_updated
  cat("**Version:**", v, "| **Owner:**", o, "| **Updated:**", u, "\n\n")
  cat("**Columns:**\n")
  for (col in ds$columns) {
    req <- if (isTRUE(col$required)) "required" else "optional"
    desc <- if (is.null(col$description)) "" else col$description
    cat("-", col$name, "(", col$type, ",", req, "):", desc, "\n")
  }
  hr <- if (is.null(ds$hard_rules)) "—" else {
    paste(vapply(ds$hard_rules, function(r) paste(names(r), unlist(r), sep = ": ", collapse = " "), character(1)), collapse = "; ")
  }
  cat("\n**Hard rules:** ", hr, "\n\n")
  sr <- ds$soft_rules
  if (!is.null(sr) && length(sr) > 0) {
    soft_txt <- vapply(sr, function(s) paste0(s$field, ": ", if (!is.null(s$message)) s$message else s$rule), character(1))
    cat("**Soft rules:** ", paste(soft_txt, collapse = "; "), "\n")
  }
} else {
  cat("Contract not available. Run pipeline first.\n")
}
```

## Validation Summary

```{r validation, results='asis'}
if (!is.na(validation$valid)) {
  status <- if (validation$valid) "PASS" else "FAIL"
  cat("**Status:** ", status, "\n\n")
  if (length(validation$errors) > 0) {
    cat("**Errors:**\n")
    for (e in validation$errors) cat("-", e, "\n")
  }
  if (length(validation$warnings) > 0) {
    cat("\n**Warnings:**\n")
    for (w in validation$warnings) cat("-", w, "\n")
  }
  if (validation$valid && length(validation$errors) == 0 && length(validation$warnings) == 0) {
    cat("No issues.\n")
  }
} else {
  cat("Validation not available. Run tar_make() first.\n")
}
```

## Metrics

### Counts by Site

```{r counts-by-site}
if (length(metrics) > 0 && !is.null(metrics$counts_by_site)) {
  knitr::kable(metrics$counts_by_site)
} else {
  cat("No metrics. Run tar_make() first.\n")
}
```

### Age Distribution Summary

```{r age-summary}
if (length(metrics) > 0 && !is.null(metrics$age_summary)) {
  knitr::kable(metrics$age_summary)
} else {
  cat("No metrics. Run tar_make() first.\n")
}
```

### Top Diagnoses

```{r top-diagnoses}
if (length(metrics) > 0 && !is.null(metrics$top_diagnoses)) {
  knitr::kable(metrics$top_diagnoses)
} else {
  cat("No metrics. Run tar_make() first.\n")
}
```

### Missingness Rates

```{r missingness}
if (length(metrics) > 0 && !is.null(metrics$missingness_rates)) {
  knitr::kable(metrics$missingness_rates)
} else {
  cat("No metrics. Run tar_make() first.\n")
}
```

## Visualizations

```{r plots, fig.width=6, fig.height=4}
fig_dir <- "output/figures"
if (dir.exists(fig_dir)) {
  for (f in list.files(fig_dir, pattern = "\\.png$", full.names = TRUE)) {
    knitr::include_graphics(f)
  }
} else {
  cat("No figures. Run tar_make() first.\n")
}
```

## Reproducibility

- **Seed:** 123 (used in `generate_synthetic_data`)
- **Package versions:** See `renv.lock` or run `renv::snapshot()` for a full lockfile.

```{r session}
sessionInfo()
```
